---
title: "Causal ML Homework 2 - code (part II)"
author: "Luis Armona and Jack Blundell"
date: "May 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r load, echo=F, message=FALSE, warning=FALSE}
# set your working directory


# clear things in RStudio

rm(list = ls())

#setwd("C:/Users/Jack/Documents/Git/AtheyMLhw2") # Jack
setwd('/home/luis/AtheyMLhw2')

# Call packages

library(ggplot2)
library(xtable)
#library(dplyr)
#library(reshape2)
#library(glmnet)
#library(plotmo)
#library(pogs)
#library(balanceHD)
library(causalTree)
library(randomForestCI)
library(reshape2)
library(plyr)
library(gradient.forest)
library(glmnet)

# set seed
set.seed(1992)

fname <- 'Data/charitable_withdummyvariables.csv'
char <- read.csv(fname)

#replace -999s with 0s (since there are already missing dummies)
for (v in names(char)){
  mi_v <- paste(v,'_missing',sep='') 
  if (mi_v %in% names(char)){
    print(paste('fixing',v))
    char[(char[,mi_v]==1),v]<-0
  }
}
```

Setup our data ready to feed into functions

```{r obs setup, echo=F, message=FALSE, warning=FALSE}

# Extract the dependent variable
Y <- char[["out_amountgive"]]

# Extract treatment
W <- char[["treatment"]]

# Extract covariates
covariates <- char[,c(14:22,23:63)] # include missing dummies for now
covariate.names <- names(covariates)

# standardize
covariates.scaled <- scale(covariates)
processed.unscaled <- data.frame(Y, W, covariates)
processed.scaled <- data.frame(Y, W, covariates.scaled)

# training, validation, and test sets.
#split evenly into 1/3,1/3,1/3 samples
set.seed(46)
sample.main <- sample(nrow(processed.scaled), round(2*nrow(processed.scaled)/3), replace=FALSE)

processed.scaled.train <- processed.scaled[sample.main,]
processed.scaled.test.C <- processed.scaled[-sample.main,]

y.train <- as.matrix(processed.scaled.train$Y, ncol=1)
y.test.C <- as.matrix(processed.scaled.test.C$Y, ncol=1)

# equal sized sets
sample.causal <- sample(nrow(processed.scaled.train), 
                     round(nrow(processed.scaled.train)/2), replace=FALSE)
processed.scaled.train.A <- processed.scaled.train[sample.causal,]
processed.scaled.train.B <- processed.scaled.train[-sample.causal,]
y.train.1 <- y.train[sample.causal]
y.train.2 <- y.train[-sample.causal]
# combination of B and C (stuff we don't directly train on)

processed.scaled.test.B.C <- rbind(processed.scaled.train.B,processed.scaled.test.C)

# write out RHS formulas for lasso heterogeneity
print(covariate.names)
sumx = paste(covariate.names, collapse = " + ")  # "X1 + X2 + X3 + ..." for substitution later
interx = paste(" (",sumx, ")^2", sep="")  # "(X1 + X2 + X3 + ...)^2" for substitution later


# Y ~ X1 + X2 + X3 + ... 
linearnotreat <- paste("Y",sumx, sep=" ~ ")
linearnotreat <- as.formula(linearnotreat)
#linearnotreat

# Y ~ W + X1 + X2 + X3 + ...
linear <- paste("Y",paste("W",sumx, sep=" + "), sep=" ~ ")
linear <- as.formula(linear)
#linear

# Y ~ W * (X1 + X2 + X3 + ...)   
# ---> X*Z means include these variables plus the interactions between them
linearhet <- paste("Y", paste("W * (", sumx, ") ", sep=""), sep=" ~ ")
linearhet <- as.formula(linearhet)
#linearhet

#calculate unconditional prop. of treatment within each sample.
processed.scaled.test.C$propens <- mean(processed.scaled.test.C$W)
processed.scaled.test.B.C$propens <- mean(processed.scaled.test.B.C$W)


# Randomized experiment. Use constant propensity score.
# Build Ystar whose Expectation=ATE to evaluate methods against

processed.scaled.test.C$Ystar <- processed.scaled.test.C$W * (processed.scaled.test.C$Y/processed.scaled.test.C$propens) -
  (1-processed.scaled.test.C$W) * (processed.scaled.test.C$Y/(1-processed.scaled.test.C$propens))


## Also do this for B.C union
processed.scaled.test.B.C$Ystar <- processed.scaled.test.B.C$W * (processed.scaled.test.B.C$Y/processed.scaled.test.B.C$propens) -
  (1-processed.scaled.test.B.C$W) * (processed.scaled.test.B.C$Y/(1-processed.scaled.test.B.C$propens))

# articificially assign B/C samples to treat/control for  ATE calculation later
processed.scaled.test.CW0 <- processed.scaled.test.C
processed.scaled.test.CW0$W <- rep(0,nrow(processed.scaled.test.C))

processed.scaled.test.CW1 <- processed.scaled.test.C
processed.scaled.test.CW1$W <- rep(1,nrow(processed.scaled.test.C))

processed.scaled.test.B.CW0 <- processed.scaled.test.B.C
processed.scaled.test.B.CW0$W <- rep(0,nrow(processed.scaled.test.B.C))

processed.scaled.test.B.CW1 <- processed.scaled.test.B.C
processed.scaled.test.B.CW1$W <- rep(1,nrow(processed.scaled.test.B.C))



# Set up MSE vector to record performance

MSElabelvec <- c("")
MSEvec <- c("")

```

Next set some parameters for causal tree / forest

```{r obs setup.params, echo=T}


# causal tree/forest params

# Set parameters
split.Rule.temp = "CT"
cv.option.temp = "CT"
split.Honest.temp = T
cv.Honest.temp = T
split.alpha.temp = .5
cv.alpha.temp = .5
split.Bucket.temp = T
bucketMax.temp= 100
bucketNum.temp = 5
minsize.temp=50


# number of trees (try 1000 once all working)
numtreesCT <- 1000
numtreesGF <- 1000


# Set parameters
split.Rule.temp = "CT"
cv.option.temp = "CT"
split.Honest.temp = T
cv.Honest.temp = T
#do regular honest weighting
split.alpha.temp = .5
cv.alpha.temp = .5
#do bucket splitting for now
split.Bucket.temp = T
bucketMax.temp= 100
bucketNum.temp = 10
minsize.temp=50

ncolx<-length(processed.scaled.train)-3 # number of possible covariates to bag from
ncov_sample<-floor(ncolx/3) #number of covariates (randomly sampled) to use to build tree
# ncov_sample<-p #use this line if all covariates need to be used in all trees


```

Use LASSO to estimate heterogeneous causal effects on Sample A

```{r obs lasso.het, echo=T}

# setup models

#create feature matrix for lasso W-int regression
#linear.train <- model.matrix(linearhet, processed.scaled.train)[,-1]
linear.test <- model.matrix(linearhet, processed.scaled.test.C)[,-1]
linear.train.1 <- model.matrix(linearhet, processed.scaled.train.A)[,-1]
linear.train.2 <- model.matrix(linearhet, processed.scaled.train.B)[,-1]

# set lower penalty factor for coef on perbush lower + interaction w/ W
p.fac = rep(1, 1 + 2*ncol(covariates))
names(p.fac) <- colnames(linear.train.1)
p.fac['W']=0
#p.fac['hpa']=0.5
#p.fac['W:hpa'] = 0.5
#p.fac['perbush']=0.5
#p.fac['W:perbush']=0.5


# Fit lasso on sample A w/ L1 penalty
set.seed(1992)
lasso.linear <- cv.glmnet(linear.train.1, y.train.1,  alpha=1, parallel=TRUE, 
                          nfolds=10,penalty.factor = p.fac)
lasso.coefs <- data.frame(coef(lasso.linear, s = "lambda.min")[which(coef(lasso.linear, s = "lambda.min") != 0)])
rownames(lasso.coefs) <- rownames(coef(lasso.linear, s = "lambda.min"))[which(coef(lasso.linear, s = "lambda.min") != 0)]
colnames(lasso.coefs) <- c("Point Estimates")
sink('Write-ups/LA/lassoint.tex')
print(xtable(lasso.coefs,caption='Lasso Heterogeneity Coefficient Estimates',label='lassoint',digits = c(4,4)))
sink()
#lasso.linear

# List non-zero coefficients
opt.lasso.coefs<-coef(lasso.linear,s='lambda.min')
selected.vars<-rownames(opt.lasso.coefs)[as.logical(opt.lasso.coefs!=0)][-1]
print(selected.vars)

# plot CV performance over complexity parameter
pdf('plots/lasso_CV.pdf')
plot(lasso.linear)
dev.off()

lasso.linear$lambda.min # min average CV error
lasso.linear$lambda.1se #  error within one s.e. of min



# set up as formula for post-selection OLS
#postOLS <- paste("Y", paste(append(selected.vars, "W"),collapse=" + "), sep = " ~ ") 
postOLS <- paste("Y ~",paste(selected.vars,collapse="+"))
postOLS <- as.formula(postOLS)

# OLS using these coefficients on sample A
lm.linear.lasso.A <- lm(postOLS, data=processed.scaled.train.A)
summary(lm.linear.lasso.A)
sink('Write-ups/LA/lassopost_A.tex')
print(xtable(summary(lm.linear.lasso.A),caption='Lasso Heterogeneity Coefficient Estimates on Sample A',label='lassointA',digits = c(4,4,4,4,4)))
sink()

# OLS using these coefficients on sample B
lm.linear.lasso.B <- lm(postOLS, data=processed.scaled.train.B)
summary(lm.linear.lasso.B)
sink('Write-ups/LA/lassopost_B.tex')
print(xtable(summary(lm.linear.lasso.B),caption='Lasso Heterogeneity Coefficient Estimates on Sample B',label='lassointB',digits = c(4,4,4,4,4)))
sink()

# OLS using these coefficients on sample C
lm.linear.lasso.C <- lm(postOLS, data=processed.scaled.test.C)
summary(lm.linear.lasso.C)
sink('Write-ups/LA/lassopost_C.tex')
print(xtable(summary(lm.linear.lasso.C),caption='Lasso Heterogeneity Coefficient Estimates on Sample C',label='lassointC',digits = c(4,4,4,4,4)))
sink()


# OLS using union of samples B and C

lm.linear.lasso.B.C <- lm(postOLS, data=processed.scaled.test.B.C)
summary(lm.linear.lasso.B.C)
sink('Write-ups/LA/lassopost_BC.tex')
print(xtable(summary(lm.linear.lasso.B.C),caption='Lasso Heterogeneity Coefficient Estimates on Samples B + C',label='lassointBC',digits = c(4,4,4,4,4)))
sink()

##############
# Predict Y on union of B and C from postOLS regression, under two counterfactual treatments, for average treatment effect
predictedW0 <- predict(lm.linear.lasso.B.C, newdata=processed.scaled.test.B.CW0)

predictedW1 <- predict(lm.linear.lasso.B.C, newdata=processed.scaled.test.B.CW1)

lassocauseff <- predictedW1-predictedW0
print(mean(lassocauseff))
# calculate MSE against Ystar (whose expectation is CATE)
lassoMSEstar <- mean((processed.scaled.test.B.C$Ystar-lassocauseff)^2)
print(c("MSE using ystar on test set of lasso",lassoMSEstar))
MSElabelvec <- append(MSElabelvec,"lasso")
MSEvec <- append(MSEvec,lassoMSEstar)


```

# Honest causal tree.

```{r hon.causal.tree, echo=T}


set.seed(1999)
CTtree <- honest.causalTree(as.formula(paste("Y~",sumx)), 
                   data=processed.scaled.train.A, treatment=processed.scaled.train.A$W, 
                   est_data = processed.scaled.train.B, est_treatment = processed.scaled.train.B$W,
                   split.Rule=split.Rule.temp,split.Honest=split.Honest.temp,
                   split.Bucket=split.Bucket.temp, bucketNum = bucketNum.temp, bucketMax = bucketMax.temp,
                   cv.option=cv.option.temp, cv.Honest=cv.Honest.temp,
                   minsize = minsize.temp, 
                   split.alpha = split.alpha.temp, cv.alpha = cv.alpha.temp, 
                   HonestSampleSize=nrow(processed.scaled.train.B))

#choose the optimal complexity penalty based on cross validation error
opcpid <- which.min(CTtree$cp[,'xerror'])
opcp <- CTtree$cp[opcpid,1]
tree_honest_CT_prune <- prune(CTtree, cp = opcp) 
#plot CV-tunes optimal honest tree
print(tree_honest_CT_prune)
pdf('./plots/honestCT.pdf')
prp(tree_honest_CT_prune)
dev.off()

# create leaf IDs for samples A,B,C up to 4 decimal places
processed.scaled.train.A$leaffact <- as.factor(round(predict(tree_honest_CT_prune, 
                                        newdata=processed.scaled.train.A,type="vector"),4))

processed.scaled.train.B$leaffact <- as.factor(round(predict(tree_honest_CT_prune, 
                                        newdata=processed.scaled.train.B,type="vector"),4))

processed.scaled.test.C$leaffact <- as.factor(round(predict(tree_honest_CT_prune, 
                                        newdata=processed.scaled.test.C,type="vector"),4))

processed.scaled.test.B.C$leaffact <- as.factor(round(predict(tree_honest_CT_prune, 
                                        newdata=processed.scaled.test.B.C,type="vector"),4))



# These show leaf treatment effects and standard errors; can test hypothesis that leaf 
# treatment effects are 0
summary(lm(Y~leaffact+W*leaffact-W-1, data=processed.scaled.train.A))
summary(lm(Y~leaffact+W*leaffact-W-1, data=processed.scaled.train.B))
summary(lm(Y~leaffact+W*leaffact-W-1, data=processed.scaled.test.C))

# Test whether each leaf treatment effects are different than ATE (uninteracted W)
summary(lm(Y~leaffact+W*leaffact-1, data=processed.scaled.test.C))
summary(lm(Y~leaffact+W*leaffact-1, data=processed.scaled.test.B.C))
sink('Write-ups/LA/leafols_test.tex')
print(xtable(summary(lm(Y~leaffact+W*leaffact-1, data=processed.scaled.test.B.C)),
             caption='OLS output of Causal Tree Leaf Regression on Left-Out Data',label='leafols',digits = c(4,4,4,4,4)))
sink()


# Predict using C (test data)
CTpredict = predict(tree_honest_CT_prune, newdata=processed.scaled.test.C, type="vector")

# calculate MSE against Ystar

CTMSEstar <- mean((processed.scaled.test.C$Ystar-CTpredict)^2)
print(c("MSE using ystar on test set of single tree",CTMSEstar))
MSElabelvec <- append(MSElabelvec,"causal tree")
MSEvec <- append(MSEvec,CTMSEstar)

```

 Now lets use causalForest from the causalTree package.
 Note: Currently CF from CT pkg gives degenerate trees; while gradient forest produces sensible expected results

```{r obs causal.f, echo=T}

# set seed
set.seed(123)
# now estimate a causalForest. Train on both A and B.
#no need to split independently A/B since the forest routine does this honest splitting for us
#rm(cf)

# fit a CT on A, estimate on B
cf.formula<-as.formula(paste("Y~",sumx))
#suppress output
log<-capture.output({
cf <- causalForest(cf.formula, data=processed.scaled.train, 
                   treatment=processed.scaled.train$W,num.trees=numtreesCT,
                   double.Sample = T, split.Bucket=T, 
                   bucketNum = bucketNum.temp,bucketMax = bucketMax.temp,
                   sample.size.total = floor(nrow(processed.scaled.train)/10),
                   sample.size.train.frac = .5,minsize = 5, 
                   cv.option=cv.option.temp, cv.Honest=cv.Honest.temp,
                   split.alpha = split.alpha.temp, cv.alpha = cv.alpha.temp,
                   split.Rule=split.Rule.temp, split.Honest=split.Honest.temp,  
                   ncolx = ncolx,ncov_sample=ncov_sample)
})

cfpredtest <- predict(cf, newdata=processed.scaled.test.C, type="vector")


cfpredtrainall <- predict(cf, newdata=processed.scaled.train.A, 
                          predict.all = TRUE, type="vector")
summary(c(cfpredtrainall$aggregate))
summary(cfpredtest)
# MSE against Ystar
cfMSEstar <- mean((processed.scaled.test.C$Ystar-cfpredtest)^2)
print(c("MSE using ystar on test set of causalTree/causalForest",cfMSEstar))
mean(cfMSEstar)

# ATE
print(c("mean of ATE treatment effect from causalForest on Training data", 
        round(mean(cfpredtrainall$aggregate),5)))

print(c("mean of ATE treatment effect from causalForest on Test data", 
        round(mean(cfpredtest),5)))


# use infJack routine from randomForestCI
# This gives variances for each of the estimated treatment effects; note tau is labelled y.hat
cfvar <- infJack(cfpredtrainall$individual, cf$inbag, calibrate = TRUE)
plot(cfvar)




```

Now lets use the gradient.forest package

```{r gradient.forest, echo=T}
##############################
# label some variables
X = as.matrix(processed.scaled.train[,covariate.names])
X.test = as.matrix(processed.scaled.test.C[,covariate.names])
Y  = as.matrix(processed.scaled.train[,"Y"])
W  = as.matrix(processed.scaled.train[,"W"])

# run gradient forest
cf.gf <- causal.forest(X, Y, W, num.trees = numtreesGF, 
                       sample.fraction=.5,
                       mtry = ncov_sample,
                       honesty=T,
                       min.node.size = minsize.temp,
                       ci.group.size = 4,
                       precompute.nuisance = FALSE,seed=1992)

# Predict on training and test sets
preds.causal.oob = predict(cf.gf, estimate.variance=TRUE)
preds.causal.test = predict(cf.gf, X.test, estimate.variance=TRUE)
summary(preds.causal.test$predictions)  
mean(preds.causal.test$predictions)  


# calculate MSE against Ystar
gfMSEstar <- mean((processed.scaled.test.C$Ystar-preds.causal.test$predictions)^2)
print(c("MSE using ystar on test set of gradient causal forest",gfMSEstar))
MSElabelvec <- append(MSElabelvec,"gradient causal forest")
MSEvec <- append(MSEvec,gfMSEstar)

# Now lets try with orthogonalization of Y (no W since random assignment)

Yrf <- regression.forest(X, Y, num.trees = numtreesGF, ci.group.size = 4)
Yresid <- Y - predict(Yrf)$prediction


cf.gfr <- causal.forest(X,Yresid,W,num.trees=numtreesGF, ci.group.size=4, 
                     precompute.nuisance = FALSE,seed=1992)
preds.causalr.oob = predict(cf.gfr, estimate.variance=TRUE)
mean(preds.causalr.oob$predictions)  
plot(preds.causalr.oob$predictions, preds.causalr.oob$variance.estimates)

Xtest = as.matrix(processed.scaled.test.C[,covariate.names])
preds.causalr.test = predict(cf.gfr, Xtest, estimate.variance=TRUE)
mean(preds.causalr.test$predictions)  
plot(preds.causalr.test$predictions, preds.causal.test$variance.estimates)


#plot heatmap of perbushXhpa ATE on residualized CFs
processed.scaled.train$leaffact <- NULL # drop this temporary var
#create fake dataset of all of the possible combos of perbush and hpa
#with medians elsewhere
namesD <- names(processed.scaled.train)
D = as.matrix(processed.scaled.train)
medians = apply(D, 2, median)

unique.hpa = sort(unique(as.numeric(D[,"hpa"])))
unique.perbush = sort(unique(as.numeric(D[,"perbush"])))
unique.vals = expand.grid(hpa = unique.hpa, perbush = unique.perbush)

D.focus = outer(rep(1, nrow(unique.vals)), medians)
D.focus[,"hpa"] = unique.vals[,"hpa"]
D.focus[,"perbush"] = unique.vals[,"perbush"]
D.focus = data.frame(D.focus[,c(-1,-2)])
numcol = ncol(D.focus)
names(D.focus) = namesD[c(-1,-2)]
df.heat = expand.grid(hpa=unique.hpa, perbush=unique.perbush)

df.heat$cate=  predict(cf.gfr, newdata=D.focus,estimate.variance=FALSE)$predictions
df.heat$hpa <- as.factor(df.heat$hpa *sd(char[,'hpa']) + mean(char[,'hpa']))
df.heat$perbush <- as.factor(df.heat$perbush *sd(char[,'perbush']) + mean(char[,'perbush']))

levels(df.heat$hpa) <- levels(as.factor(round(as.numeric(as.character(df.heat$hpa)),2)))
levels(df.heat$perbush) <- levels(as.factor(round(as.numeric(as.character(df.heat$perbush)),4)))
#ggplot(df.heat, aes(x=hpa,y=perbush,fill=cate,colour=cate)) + geom_tile() +
#  scale_fill_gradient(low = "red", high = "blue") + scale_colour_gradient(low = "red", high = "blue")

#ggplot(df.heat, aes(x=hpa,y=perbush,fill=cate)) + geom_raster(interpolate=F) 
gg<-ggplot(df.heat, aes(x=hpa,y=perbush,fill=cate,colour=cate)) + geom_tile() +
  scale_x_discrete(breaks = levels(factor(df.heat$hpa))[c(T, rep(F, 14))]) +
  scale_y_discrete(breaks = levels(factor(df.heat$perbush))[c(T, rep(F, 4))]) + 
  ylab("% voted for Bush in '04 ")  + xlab('Previous contribution ($)') +
  scale_fill_gradient(low = "red", high = "blue") + scale_colour_gradient(low = "red", high = "blue")
print(gg)
ggsave('plots/heatmap.pdf')

#plot hist of resid GF hists
gg<-ggplot(data.frame(preds.causalr.test),aes(x=predictions,fill=I('white'),col=I('black')) ) + geom_density()
print(gg)
ggsave('plots/cate_gf.pdf')
# calculate MSE against Ystar
gfrMSEstar <- mean((processed.scaled.test.C$Ystar-preds.causalr.test$predictions)^2)
print(c("MSE using ystar on test set of orth causal gradient forest",gfrMSEstar))
MSElabelvec <- append(MSElabelvec,"orth causal gradient forest")
MSEvec <- append(MSEvec,gfrMSEstar)

pdf('./plots/residGFcomp_random.pdf')
plot(preds.causal.test$predictions,preds.causalr.test$predictions,
         xlab='GF CATEs (no residualization)',ylab='GF CATEs (pre-residualized)')
abline(0,1)
dev.off()

#print final output
#print(MSElabelvec)
names(MSEvec) <- MSElabelvec
print(MSEvec)
```

